---
title: "EDS 223 Homework Assignment 4"
subtitle: "Prioritizing potential aquaculture"
author: "Rachel Swick"
date: last-modified
execute:
  warning: false
  message: false
format:
  html:
    toc: true
    code-fold: true
editor_options: 
  chunk_output_type: console
---

## Project Background

Marine aquaculture has the potential to play an important role in the global food supply as a more sustainable protein option than land-based meat production.1 Gentry et al. mapped the potential for marine aquaculture globally based on multiple constraints, including ship traffic, dissolved oxygen, and bottom depth. They found that global seafood demand could be met using less than 0.015% of the global ocean area.2

## Learning Goals

This assignment will reinforce key concepts in geospatial analysis by practicing the following:

- combining vector/raster data
- resampling raster data
- masking raster data
- map algebra

## About the Data

## Load Packages and Import Data
```{r}
# Load packages
library(terra)
library(stars)
library(tidyverse)
library(tmap)
library(kableExtra) 
library(sf)
library(here)
library(testthat)
library(devtools)
library(remotes)
library(patchwork)
```

```{r}
# West Coast EEZ Data
west_coast_eez <- read_sf(here("data", "wc_regions_clean.shp"))

# Bathymetry (depth) Data
bathymetry <- rast("data/depth.tif")

# Average Annual SST Data for 2008 to 2012
sst_2008 <- rast("data/average_annual_sst_2008.tif")
sst_2009 <- rast("data/average_annual_sst_2009.tif")
sst_2010 <- rast("data/average_annual_sst_2010.tif")
sst_2011 <- rast("data/average_annual_sst_2011.tif")
sst_2012 <- rast("data/average_annual_sst_2012.tif")
```

```{r}
# Combine SST data into a raster stack
sst_stack <- merge(sst_2008, sst_2009, sst_2010, sst_2011, sst_2012)
```

## Ensure Matching Coordinate Reference Systems (CRS)

An accurate map can not be created without first ensuring all spatial objects are using the same CRS.

```{r}
# Create list of all spatial objects
spatial_objects <- list(west_coast_eez,
                        bathymetry,
                        sst_stack)

# Check if all spatial objects have the same CRS
# If not, determine whether it is a vector or raster object
# Update CRS based on object type
for (i in 2:length(spatial_objects)) {
  if (inherits(spatial_objects[[i]], "SpatRaster") && 
      !identical(crs(spatial_objects[[i]]), spatial_objects[[1]])) {
    spatial_objects[[i]] <- project(spatial_objects[[i]], crs(spatial_objects[[1]]))
    warning("The CRS has been transformed to match.")
  }
  else if (inherits(spatial_objects[[i]], c("sf", "tbl_df", "tbl", "data.frame")) &&
      st_crs(spatial_objects[[i]]) != st_crs(spatial_objects[[1]])) {
    spatial_objects[[i]] <- st_transform(spatial_objects[[i]], st_crs(spatial_objects[[1]]))
    warning("The CRS has been transformed to match.")
  }
  else {
    print("The CRS Match.")
  }
}

# Check that all spatial objects have been transformed
for (i in 2:length(spatial_objects)) {
  print(st_crs(spatial_objects[[i]]) == st_crs(spatial_objects[[1]]))
}
```

## Process SST and Depth Data

Process the SST and depth data so that they can be combined into a single raster object. This is necessary because the SST and depth data have slightly different resolutions, extents, and positions.

```{r}
# Mean SST from 2008 to 2012
mean_sst <- mean(sst_stack, na.rm = TRUE)

# Convert mean SST from Kelvin to Celsius
mean_sst_c <- mean_sst - 273.15

# Crop depth raster to the extent of average sst raster
bathymetry_crop <- crop(bathymetry, mean_sst_c) 

# Resample depth data to match the resolution of the average sst data
bathymetry_resample <- resample(bathymetry_crop, y = mean_sst_c, method = "near")

# Combine depth and average sst
sst_depth_stack <- merge(bathymetry_resample, mean_sst_c)
```

## Find Suitable Locations

```{r}

```

























































